---
title: "Text analytics for large sets of customer feedback"
author: "phi.designer"
date: "11 August 2015"
output: ioslides_presentation
incremental: true
smaller: true
logo: noun_168650_cc.png
---

## Objectives

According to IBM, about *"80% of the information created and used by an enterprise is unstructured data, and this figure is growing at twice the rate of structured data”*. 

Hence, it becomes imperative to **developed the required skills and tools to unlock this vast source of information** for critical business insights.

> - The [accompanying app](https://phidesigner.shinyapps.io/Final_app) for this presentation, aims at exposing users to some of the most common, yet simple, techniques used for text mining, in an **interactive and simple way**.

Further information on text mining can be gathered from Yanchang Z. [Text Mining with R tutorials](http://www.rdatamining.com/examples/text-mining).

## Methodology

The [app](https://phidesigner.shinyapps.io/Final_app) exposes users to simple text mining techniques, by allowing the interaction with three common Text Mining tools: **Word counting, Tree clustering and Kmeans analysis for word proximity**, through the following interactions:

> - **1) Bar chart for word counting across all feedback:** Choose lower percentage threshold of word usage across all feedback (e.g. words used on at least 50 % of all feedback).

> - **2) Dendogram of feedback’s natural clustering:** Choose max allowed sparsity threshold (i.e. resulting cluster contains only terms with a sparse factor of less than sparse value)

> - **3) Verbatim of clusters by word proximity (K means):** Choose number of words per cluster, so as to better understand narrative.

## Dendogram chart example | Promoter's feedback {.flexbox .vcenter}

```{r, echo=FALSE, cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
# requried libraries
library(tm)
library(SnowballC)
library(ggplot2)
library(ggdendro)

# reading file
data <- read.csv("Book1.csv", stringsAsFactors = TRUE)
Promo <- as.data.frame(data[data$Score == "Prom", 2]); names(Promo) <- "Reason"

# creating corpus
# Promo
Corpus_Promo <- Corpus(VectorSource(Promo$Reason))
Corpus_Promo <- tm_map(Corpus_Promo, content_transformer(tolower))
Corpus_Promo <- tm_map(Corpus_Promo, removePunctuation)
Corpus_Promo <- tm_map(Corpus_Promo, removeWords, stopwords("english"))

# Freq matrix (words >3 characters)
tdm_Promo <- TermDocumentMatrix(Corpus_Promo, control = list(wordLengths = c(3, Inf)))

# Check for sparsity (most Freq terms)
frq_Promo <- findFreqTerms(tdm_Promo, lowfreq=50)

# Check for sparsity
# Promo
term.frq_P <- rowSums(as.matrix(tdm_Promo))
term.frq_P <- subset(term.frq_P, term.frq_P >= 50)
df_P <- data.frame(term = names(term.frq_P), freq = term.frq_P)

# Clustering
# remove sparse terms
Promo_clust <- removeSparseTerms(tdm_Promo, sparse = 0.93)
Promo_matrix <- as.matrix(Promo_clust)

# Plotting cluster terms
dist_Promo <- dist(scale(Promo_matrix)) # clustering terms by proximity
fit_Promo <- hclust(dist_Promo, method = "ward.D2")
dendr    <- dendro_data(fit_Promo, type="rectangle")
clust    <- cutree(fit_Promo,k=4)
clust.df <- data.frame(label=names(clust), cluster=factor(clust))
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
ggplot() + geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=label(dendr), aes(x, y, label=label, hjust=0, color=cluster), 
            size=4) + theme_minimal() + coord_flip() + scale_y_reverse(expand=c(.2,0)) +
               labs(title = "Dendogram: Clustering by themes", x = "Themes", y = "Distance")
```

## Dendogram's code {.smaller}

```{r, echo=TRUE, eval=FALSE}
# Remove sparse terms
Promo_clust <- removeSparseTerms(tdm_Promo, sparse = 0.93)
Promo_matrix <- as.matrix(Promo_clust)
# Plotting cluster terms
dist_Promo <- dist(scale(Promo_matrix))
fit_Promo <- hclust(dist_Promo, method = "ward.D2")
dendr    <- dendro_data(fit_Promo, type="rectangle")
clust    <- cutree(fit_Promo,k=4)
clust.df <- data.frame(label=names(clust), cluster=factor(clust))
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
# Dendogram plot
ggplot() + geom_segment(data=segment(dendr), aes(x=x, y=y,
          xend=xend, yend=yend)) + geom_text(data=label(dendr), 
          aes(x, y, label=label, hjust=0, color=cluster), 
          size=4) + theme_minimal() + coord_flip() 
          + scale_y_reverse(expand=c(.2,0)) 
          + labs(title = "Dendogram: Clustering by themes", 
          x = "Themes", y = "Distance")
```

... Thanks!

PS: This presentation was done in RStudio presenter using "output: ioslides_presentation" for the Data Science Specialisation (JHU/Coursera)